---
title: "Shiny Integration"
output:
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 3
vignette: >
  %\VignetteIndexEntry{Shiny Integration}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
editor_options:
  markdown:
    wrap: 72
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>",
  eval     = FALSE
)
```

## Introduction

`gridify()` always returns a **grid graphical object**, regardless of whether
the input is a figure or a table. It must therefore always be displayed using
`plotOutput` / `renderPlot` in a Shiny application.

The `gridify_with_settings` module wraps this requirement in a reusable,
self-contained Shiny module that works with any `gridify` output type:
`ggplot2` figures, base R figures, `flextable`, `gt`, and `rtables`.

---

## The `gridify_with_settings` module

The module consists of two functions:

- **`gridify_with_settings_ui(id)`** — the UI side, built as a
  `sidebarLayout` with sliders and download buttons on the left and a
  dynamic plot panel on the right.

- **`gridify_with_settings_srv(id, gridify_r, height, width)`** — the
  server side. `gridify_r` must be a `shiny::reactive()` that returns a
  `gridifyClass` object (the result of `gridify()`). `height` and
  `width` each take a `numeric(3)` vector of `c(value, min, max)`;
  defaults are `c(600L, 200L, 2000L)` and `c(800L, 200L, 2000L)`
  respectively.

---

## Example 1 — ggplot2 figure with `pharma_layout_base()`

This mirrors the pharma layout example shown in the
[Simple Examples](simple_examples.html) vignette, wrapped in a Shiny
app. See that vignette for the full list of available cells and layout
options.

```{r ggplot_shiny}
library(shiny)
library(ggplot2)
library(gridify)

ui <- fluidPage(
  titlePanel("gridify + Shiny: ggplot2 figure"),
  gridify_with_settings_ui("gg_demo")
)

server <- function(input, output, session) {
  gridify_r <- reactive({
    gridify(
      object = ggplot2::ggplot(mtcars, aes(x = mpg, y = wt)) +
        ggplot2::geom_point(),
      layout = pharma_layout_base()
    ) |>
      set_cell("header_left_1", "My Company") |>
      set_cell("header_left_2", "PROJECT / INDICATION") |>
      set_cell("header_left_3", "STUDY") |>
      set_cell("title_1", "Weight vs Fuel Efficiency") |>
      set_cell("footer_left", paste("Program: my_script.R,", Sys.Date())) |>
      set_cell("footer_right", "Page 1 of 1") |>
      set_cell("watermark", "DRAFT")
  })

  gridify_with_settings_srv("gg_demo", gridify_r = gridify_r)
}

shinyApp(ui, server)
```

---

## Example 2 — `rtable` table

`gridify` supports `rtables` objects directly — the conversion is handled internally.
Pass the `rtable` straight to `gridify()` just like any other object type.

> **Note:** the internal conversion uses `rtables.officer::tt_to_flextable()` with
> default settings. If you need to customise the font, borders, or cell styling, convert
> manually to a `flextable` first (see the [Simple Examples](simple_examples.html)
> vignette for the step-by-step approach).

```{r rtable_shiny}
library(shiny)
library(gridify)
library(rtables)  # install.packages("rtables")

ui <- fluidPage(
  titlePanel("gridify + Shiny: rtable"),
  gridify_with_settings_ui("rt_demo")
)

server <- function(input, output, session) {
  gridify_r <- reactive({

    # Build the rtable
    tbl <- rtables::basic_table(main_footer = " ") |>
      rtables::split_cols_by("Species") |>
      rtables::analyze(
        c("Sepal.Length", "Sepal.Width", "Petal.Length"),
        function(x, ...) {
          rtables::in_rows(
            "Mean (sd)" = c(mean(x), sd(x)),
            "Median"    = median(x),
            "Min - Max" = range(x),
            .formats = c("xx.xx (xx.xx)", "xx.xx", "xx.xx - xx.xx")
          )
        }
      ) |>
      rtables::build_table(iris)

    # Pass directly to gridify — no manual flextable conversion needed
    gridify(
      object = tbl,
      layout = pharma_layout_base()
    ) |>
      set_cell("header_left_1", "My Company") |>
      set_cell("header_left_2", "PROJECT / INDICATION") |>
      set_cell("header_left_3", "STUDY") |>
      set_cell("title_1", "Summary Statistics for Iris Dataset") |>
      set_cell("footer_left", paste("Program: my_script.R,", Sys.Date())) |>
      set_cell("footer_right", "Page 1 of 1") |>
      set_cell("watermark", "DRAFT")
  })

  gridify_with_settings_srv("rt_demo", gridify_r = gridify_r)
}

shinyApp(ui, server)
```
